Alleviating the Topology Mismatch Problem in Unstructured and Structured Overlay Networks: A Survey

### 2014-01-21 mard-survey.tex
2012.03.23 (Vassilis) Adding an abstract. Introduction rewritten to use first person active voice. Introduction restructured for better coherence. Better integration of the ACM LaTeX class. Bibliography not rendered is a know issue.

%\documentclass[a4paper,10pt]{article}
\documentclass[acmcsur]{acmtrans2m}
\usepackage[utf-8]{inputenc}

\usepackage{paralist}

\usepackage{graphicx}

\usepackage{colortbl}
\usepackage{color,soul}
\usepackage{longtable}
\usepackage{pdflscape}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{definition}[theorem]{Definition}

@@

%opening
% TODO: Contribution of this paper
% Q&A: 'Overlay' or 'Peer-to-Peer'/'P2P'
\title{Alleviating the Topology Mismatch Problem in Unstructured and Structured Overlay Networks: A Survey}
\author{
Vassilis Moustakas$^1$, H\"useyin Akcan$^2$, Mema Roussopoulou$^1$ and Alex Delis$^1$\\
$^1$Department of Informatics and Telecommunications,\\
National and Kapodistrian University of Athens\\
\mbox{\texttt{\{b.moystakas, mema, ad\}@di.uoa.gr}}\\
$^2$Department of Software Engineering,\\
Izmir University of Economics, Izmir, Turkey \\
\mbox{\texttt{huseyin.akcan@ieu.edu.tr}}
}

\begin{document}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
% TODO: Contribution of this paper
Peer-to-peer (p2p) systems is a rapidly growing application architecture. The overlay network abstraction on top of a best-effort infrastructure provides a powerful tool to realize ever-wanted exotic features like anonymity, high availability and robustness, load balancing, QoS at the application layer and many more. Unfortunately weaknesses of early effectuations constrained applications from unleashing the full potential of the paradigm. One of them is the topology mismatch problem between the overlay and the physical underlying networks, a problem that imposes unnecessary stress to the network resources. This survey investigates on the recent research developments towards the alleviation of the problem, in both unstructured and structures p2p schemes.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

% TODO: Contribution of this survey
% UNDER CONSTRUCTION

% TODO: Structure of this paper
% UNDER CONSTRUCTION!
This survey begins with a quick review of the conception of the overlay network and the available architectures of p2p systems. It continues with the identification of the phenomenon of the topology mismatch between overlay and underlying networks and its impact on the utilization of network resources. Ultimately, it reviews the most important resources found in the literature that attempt to tackle the problem as it proposes a taxonomic scheme, using a set of supertype-subtype relationships based on unique characteristics they have and/or specific goals they target.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% OK
\section{Overlay Networks and P2P Systems}

\begin{figure}[h!]
\centering
  \includegraphics[scale=0.5]{img/p2p.pdf}
\caption{An example overlay network.}
\label{figure:overlay}
\end{figure}

An \emph{overlay network} is an abstract, logical interconnection of entities
formed on top of another network. It consists of a set of nodes, the computing
elements, that are being connected by virtual (logical) point-to-point links.
For example, the \emph{Asymmetric Digital Subscriber Line (ADSL)} is, actually,
an overlay network on top of the \emph{Public Switched Telephone Network
(PSTN)}. Figure \ref{figure:overlay} illustrates an example physical ethernet network
and the corresponding overlay network formed by four nodes. 
Nodes $A$ and $B$ are on the same local network, while $C$ and $D$ are in
different networks. The top layer represents the overlay network formed by these
nodes, which is a reflection of the application layer connections among these
nodes. As the topology of the network can change based on the application, in
this example, it is easy to observe that even though nodes $A$ and $B$ are in the same
network, their communication goes through node $C$ on another network,
demonstrating a non-optimal mapping between the physical and the overlay
networks.
An overlay network is generally useful when
abstraction of the network layer from the application layer is important for
applications, particularly when the underlying network structure is subject to
frequent changes.%, which is a common case for ad-hoc networks.

The abstraction of overlay networks has been proposed as a way to implement
efficient, fully distributed, application layer services on top of a best-effort
IP layer forming a widely ranged family of protocols collectively known as
\emph{peer-to-peer} or \emph{p2p} for short. P2p network systems gained
significant attention, in recent years, due to their plethora of unique features
in supporting file sharing among huge numbers of (inter-) networked computers
called \emph{peers}, where each, such, \emph{peer} could act both as a resource
provider and a resource consumer. This changed the traditional
\emph{client-server} model dominating the internet and lead to the introduction
of the \emph{servent}-concept \cite{gnutella}, a portmanteau that blends the
notions of \emph{server} and \emph{client} to denote the twofold role of the
participants in a p2p network. Such serverless systems, proved to be able to
achieve outstanding aggregate resource capacities as participants join the
system
%\footnote{Unfortunatelly, there has been observed mitigation of this
%self-scaling property by the undesirable behaviour of participating users,
%usually answered in popular file-sharing systems; from \emph{free-riding} (i.e.
%in Napster and Gnutella networks \cite{saroiu_measurefileshare_2002,
%adar_gnutellafreeriders_2000, hughes_gnutellafreeride_2005}), to the
%distribution of illegal content and/or other \emph{socio-technical} issues
%\cite{hughes_socp2p_2008}.}, 
without requiring additional expenditure for infrastructure.

As p2p systems evolved, three main architectures have been emerged, namely:
\begin{itemize}
  \item \emph{centralized}
  \item \emph{decentralized unstructured}
  \item \emph{decentralized structured}
\end{itemize}

\emph{Centralized} architectures were the first to recognize that requests for popular content need not be sent to a central server but instead could be handled by the many hosts that already posses the content. \emph{Napster}, for example, maintained a \emph{central index server} based on file lists provided by participating peers. The central index server was queried by the users and it returned pointers to the actual content. Thus, by centralizing search while distributing downloads, Napster achieved a highly functional design that was widely acknowledged, at the time, as ``the fastest growing Internet application ever''. This centralized search facility, though, renders this scheme not fully distributed and can be proven its ``Achilles' heel'' in terms of being a scalability bottleneck, a central point of failure and vulnerable to malicious acts (e.g. \emph{Denial-of-Service (DoS) attacks}).

The centralized approach has been eventually replaced by architectures that distributed both the search and the download capabilities. In these \emph{decentralized unstructured} architectures, file placement is random, which means there is no correlation with the network topology whatsoever \cite{yang_improvep2psearch_2002}. The most important properties of such systems are that they support inherent heterogeneity of peers, are highly resilient to peer failures, and incur low maintenance overhead at handling the dynamics of peer participation \cite{stutzbach_churn_2006}. In the literature, they are also known as \emph{broadcast-based} systems, because they use \emph{message flooding} among peers (i.e. Gnutella\cite{gnutella}) or among super-peers (i.e KaZaA\cite{kazaa}) to propagate queries. 
Even though the unstructured approach became highly popular among file sharing applications, they do have a certain disadvantage in locating rare objects due to the unstructured nature and the limitations of the flooding approach. 
%Each (super-) peer receiving a query decreases the incoming query message's lifetime which is measured in \emph{time-to-live (TTL)} hops, back-propagates its results, and  if the lifetime of the incoming message did not expired (TTL did not reach zero), it forwards the message to all its neighbours except from the one it receives it from.
% TODO: maybe enumerate applications of decentralized unstructured p2ps.

Recently, \emph{decentralized structured} schemes have been proposed in order to provide a self-organising infrastructure for large-scale p2p applications \cite{ratnasamy_can_2001,stoica_chord_2001,antony_pastry_2001,zhao_tapestry_2001,maymounkov_kademlia_2002}. They implement a \emph{Distributed Hash Table (DHT)} that maps objects to nodes through a deterministic mechanism. The main advantage of the DHT approach of decentralized structured p2p networks over decentralized unstructured ones, is that they provide a guaranteed bound on the number of overlay routing hops that have to be taken in order for an object to be located even in the case where only a single copy exists in the system. This is $O \left ( log n \right )$ compared to Gnutella, that requires $O \left ( n \right )$ to reliably locate a specific object.
% TODO: maybe enumerate applications of decentralized structured p2ps. (OceanStore, Far)

%{\sethlcolor{yellow}\hl{HA: Give the brief advantages and disadvantages of structured and unstructured
%approaches here!}}

% TODO: enumerate applications of overlay networks and p2p systems regardless of architecture
% UNDER CONSTRUCTION
% Q&A: Need more??
Initial efforts targeted on the implementation of one-to-many addressing schemes
that could replace \emph{IP multicast} in providing higher quality of streaming
media through \emph{quality of service (QoS)} guarantees. IP multicast as well
as other proposals such as \emph{IntServ} and \emph{DiffServ}
\cite{cisco_diffserv_2005} have not seen wide acceptance (yet?), largely because
they require changes to all routers of the network. On the other hand, an
overlay network can be incrementally deployed on end-systems running the overlay
protocol software, without cooperation from the ISPs. Academic research includes
\cite{chu_esm_2000,jannotti_overcast_2000,kwon_tag_2002} among others.  Overlay
networks were also implemented in order to back the routing of messages to
destinations whose address is not known in advance. This resulted in the
appearance of well known peer-to-peer protocols such as \cite{gnutella} and
\cite{maymounkov_kademlia_2002} widely used for digital content sharing among
their network nodes.  Other special purpose overlay networks include
\cite{anderson_ron_2001} for resilient routing; \cite{subramanian_overqos_2004}
for quality of service guarantees; and \cite{clarke_freenet_2001} for anonymity;
to name just a few.

The main focus of this paper is the topology mismatch problem, therefore the
available approaches in overlay networks are briefly summarized, and interested
readers are pointed to the available surveys on overlay networks
\cite{TheotokisS04,LuaCPSL05}.

>>>>

%\documentclass[a4paper,10pt]{article}
\documentclass[acmcsur,acmnow]{acmtrans2m}
%\documentclass[acmtocl,acmnow]{acmtrans2m}

%\acmVolume{V}
%\acmNumber{N}
%\acmYear{YY}
%\acmMonth{Month}

% packages
\usepackage[utf-8]{inputenc}
\usepackage{paralist}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{colortbl}
\usepackage{color,soul}
\usepackage{longtable}
\usepackage{pdflscape}

% definitions
\newtheorem{theorem}{Theorem}[section]
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newdef{definition}[theorem]{Definition}
\newdef{remark}[theorem]{Remark}

%\newtheorem{theorem}{Theorem}[section]
%\newtheorem{definition}[theorem]{Definition}

@@

\markboth{}{}{}

\title{Alleviating the Topology Mismatch Problem in Unstructured and Structured Overlay Networks: A Survey}

\author{
VASSILIS S. MOUSTAKAS\\\emph{National and Kapodistrian University of Athens}\\
H\"USEYIN AKCAN\\\emph{Izmir University of Economics}\\
MEMA ROUSSOPOULOS\\\emph{National and Kapodistrian University of Athens}\\
AND\\
ALEX DELIS\\\emph{National and Kapodistrian University of Athens}\\
%  Vassilis Moustakas$^1$, H\"useyin Akcan$^2$, Mema Roussopoulou$^1$ and Alex Delis$^1$\\
%  $^1$Department of Informatics and Telecommunications,\\
%  National and Kapodistrian University of Athens\\
%  \mbox{\texttt{\{b.moystakas, mema, ad\}@di.uoa.gr}}\\
%  $^2$Department of Software Engineering,\\
%  Izmir University of Economics, Izmir, Turkey \\
%  \mbox{\texttt{huseyin.akcan@ieu.edu.tr}}
}

\begin{abstract}
Peer-to-peer (P2P) systems is a rapidly growing application architecture. The overlay network abstraction on top of a best-effort infrastructure provides a powerful tool to realize a plethora of ever-wanted exotic features like anonymity, high availability and robustness, load balancing, quality-of-service (QoS) at the application layer and many more. Unfortunately weaknesses of early effectuations constrained applications from unleashing the full potential of the paradigm. One of them is the topology mismatch problem between the overlay and the physical underlying network, a problem that imposes a huge amount of unnecessary stress to the network resources and, thus, impede applications from scaling. In this survey we investigate on the recent research developments towards the alleviation of this problem, in both unstructured and structured overlay architectures. We visit the various algorithms available in the lietrature, focusing on the presentation of how each of them tries to tackle the problem. Additionally, we provide a finer grained categorisation of the approaches based on the generic type of methodology they use in order to accomplish this goal.
% TODO: Review this after the comparison sections are fixed.
Finally we discuss the advantages, disadvantages and novelties introduced by the algotithms and offer pictorial comparisons for them in order to further strengthen and organise understanding of the information and knowledge cited, in an attempt to make this survey a one-stop-shop reference for anyone interested in the topic, from researchers and students, to system architects and developers.
\end{abstract}

\category{}{}{}

\terms{Topology mismatch, distributed, peer-to-peer, P2P, overlay}

\keywords{Topology mismatch problem, distributed algorithm, peer-to-peer architecture, overlay network }

\begin{document}

\begin{bottomstuff}
SAY SOMETHING COOL HERE...
\end{bottomstuff}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Distributed architectures have witnessed a significant increase of interest during the last decade, as the number of interconnected computers increased, the available digital information kept proliferating, at an exponetial rate each year, and problems to be addressed by IT became more and more demanding. Why is there a correlation, and it is not just a coincidence? Distributed computation suggests a simple paradigm\footnote{Of course, as always, this is the case only after someone thought about it first!}. Get a set of autonomous computing entities and put them to cooperate with each other in order to fullfil a well defined task. In the World of the Internet and the World Wide Web this also offers a thrilling promise. To harness the immence power scattered across the Internet and get amply rewarded by exploiting the data, storage and computational resources out there.

In order to achieve that, computers collaborate by exchanging messages via an application-level network where they can address each other in the specific context of the task at hand. This communication structure is refered to as an \emph{overlay network} which is an abstract, logical inter-connection formed on top of another network\footnote{The Internet itself is an overalay on top of the telephone network.}. It consists of a set of nodes, the computing elements, that are being connected by virtual (logical) point-to-point links, perhaps corresponding to multiple physical links. Figure~\ref{figure:overlay} illustrates a simple four-node overlay contructed upon a wide area network.

\begin{figure}
\centering
  \includegraphics[scale=0.5]{img/p2p.pdf}
\caption{An example overlay network.}
\label{figure:overlay}
\end{figure}

The abstraction of the overlay network has been proposed as a way to implement efficient, fully distributed, application-layer services. For example, backing the routing of messages to destinations that are not known in advance or building one-to-many addressing schemes
% TODO find citation
%[Narada, Adobe's Real Time Media Flow Protocol (RTMFP)]
that would provide higher quality-of-service (QoS) guarantees. \emph{IP multicast}, as well as other proposals such as \emph{IntServ} and \emph{DiffServ} \cite{cisco_diffserv_2005}, have not seen wide acceptance (yet?), largely due to the fact they require support from the physical infrastructure. On the other hand, an overlay network can be deployed on end-systems running the overlay protocol software, without cooperation from the ISPs.
%Academic research includes \cite{chu_esm_2000,jannotti_overcast_2000,kwon_tag_2002} among others.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Peer-to-Peer (P2P) Architectures in a Nutshell}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Overlay networks on top of the best-effort IP layer, formed a widely ranged family of protocols collectively known as \emph{peer-to-peer} or \emph{P2P} for short. P2P systems consist of a number of networked computers called \emph{peers}, where each, such, \emph{peer} could act both as a resource provider and a resource consumer. This changed the traditional \emph{client-server} model dominating the Internet and lead to the introduction of the \emph{servent} concept \cite{gnutella}, a \emph{portmanteau} that blends the notions of server and client to denote the twofold role of the P2P network participants. Such serverless systems, proved to be able to achieve outstanding aggregate resource capacities as more and more participants join the system without requiring additional expenditure for infrastructure\footnote{
        Unfortunatelly, there has been observed mitigation of this self-scaling property by the undesirable behaviour of participating users, usually answered in popular file-sharing systems; from \emph{free-riding} \cite{saroiu_measurefileshare_2002, adar_gnutellafreeriders_2000, hughes_gnutellafreeride_2005} to the distribution of illegal content and/or other \emph{socio-technical} issues \cite{hughes_socp2p_2008}.
}.
P2p network systems gained extra attention whithin the distributed systems super-set, circa 2000, due to the fact that they succesfully supported popular applications that offered file sharing among vast numbers of participating users. But their development didn't stay still. On the contrary, during the last decade, research on P2P systems was constant and fruitfull generating a wide range of popular protocols, networks and applications. As P2P systems evolved three main architectures have been emerged, namely
\begin{inparaenum}[\itshape i\upshape)]
  \item \emph{centralized},
  \item \emph{decentralized unstructured}, and
  \item \emph{decentralized structured}.
\end{inparaenum}

\emph{Centralized} architectures (see Figure~\ref{figure:p2p-archs:centralised}) were the first to recognize that requests for resourses (e.g. CPU cycles \cite{seti} or popular content 
% TODO find citation for Napster
\cite{}) need not be sent to a dedicated server but instead could be handled by the many hosts that already posses them. This schema, though, is not fully distributed since it dictates a centralised searching facility, a characteristic that can be proven its ``Achilles' heel'' in terms of being a scalability bottleneck, a central point of failure and vulnerable to malicious acts (e.g. \emph{Denial-of-Service (DoS) attacks}).

The centralized approach has been eventually replaced by architectures that distributed both search and download capabilities (see Figures~\ref{figure:p2p-archs:unstructured-full} and \ref{figure:p2p-archs:unstructured-hierarchical}). In these \emph{decentralized} architectures, file placement is random, which means there is no correlation with the network topology whatsoever \cite{yang_improvep2psearch_2002}. For this reason they are also refered to as \emph{unstructured}. The most important properties of such systems are that they support inherent heterogeneity of peers, are highly resilient to peer failures, and incur low maintenance overhead at handling the dynamics of peer participation \cite{stutzbach_churn_2006}. In the literature, they are also known as \emph{broadcast-based} systems, because they use \emph{message flooding} among peers or super-peers to propagate queries
%\footnote{The life of each query message is constrained by an integer, refered to as \emph{time-to-live (TTL)}. When a querry is received by a (super-) peer, this value is decreased by one unit. Then, the receiver, back-propagates its results and if the TTL has not reached zero the node forwards the query message to all its neighbours except from the one it received it from.}
.

Recently, \emph{decentralized structured} schemes have been proposed (see Figure~\ref{figure:p2p-archs:structured}) in order to provide a self-organising infrastructure for large-scale P2P applications \cite{ratnasamy_can_2001, stoica_chord_2001, antony_pastry_2001, zhao_tapestry_2001, maymounkov_kademlia_2002, rgrk_bamboo_2004}. They implement a \emph{Distributed Hash Table (DHT)} that maps objects to nodes through a deterministic mechanism. Even though unstructured networks became highly popular among file sharing applications, they do have a certain limitation in locating rare objects. The DHT approach of structured networks on the other hand, provides a guaranteed bound on the number of overlay routing hops that have to be taken in order for an object to be located, even in the case where only a single copy exists in the system.  This means $O \left ( log n \right )$ hops, compared to unstructured networks that require $O \left ( n \right )$ to reliably locate a specific object\footnote{Unfortunatelly this efficiency does not come without cost. Structured overlays realise what is called object identifier location as opposed to content-based retrieval of unstructured ones. This means that the later can support more versatile object location mechanisms like keyword and partial matching searches.}.

% TODO: Change these pictures to match look-n-feel
\begin{figure}[ht]
\centering
\subfigure[Centralised (SETI@Home or Napster).] {
  \includegraphics[scale=0.3]{img/centralised.png}
  \label{figure:p2p-archs:centralised}
}\qquad\qquad
\subfigure[Fully unstructured (Gnutella v.0.4).] {
  \includegraphics[scale=0.3]{img/unstructured-full.png}
  \label{figure:p2p-archs:unstructured-full}
}\qquad\qquad
\subfigure[Hierarchical unstructured (Gnutella v.0.6 or FastTrack).] {
  \includegraphics[scale=0.3]{img/unstructured-hierarchical.png}
  \label{figure:p2p-archs:unstructured-hierarchical}
}\qquad\qquad
\subfigure[Structured (Chord and Kademlia).] {
  \includegraphics[scale=0.3]{img/structured.png}
  \label{figure:p2p-archs:structured}
}
\caption{The various peer-to-peer architectures.}
\label{figure:p2p-archs}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Applications and Implications}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The most successful incarnation of the centrilised approach is without a doubt the popular file-sharing application \emph{Napster}\footnote{Pioneering, though, was the famous \emph{SETI@Home Project} \cite{seti}, which pools spare processing power of participants to search the universe for the existence of extra terrestrial intelligence!}
% TODO find citation for Napster again
\cite{}.
Napster maintained a \emph{central index server} based on file lists provided by participating peers. The central index server was queried by the users and it returned pointers to the actual content. Thus, by centralizing search while distributing downloads, Napster achieved a highly functional design that was widely acknowledged, at the time, as ``the fastest growing Internet application ever'' that toped some 26 million users and 80 million songs \cite{jmm_naptopusage_2001}.

Napster was ultimatelly brought down after Recording Industry Association of America (RIAA) filed a lawsuit for copyright infringement of the intellectual property of artists whose music was distributed by the network\footnote{
      The sueing spree was initiated by heavy metal giants \emph{Metallica} that feared Napster was undermining record sales. Opposed opinions instantly emerged  with best example the british alternative rock band \emph{Radiohead} whose album \emph{Kid A} had also been leaked to Napster three months before reaching music store shelves \cite{rm_radioheadkida_2000}. Despite music critics labeling the album ``experimental'' and ``challenging'' (it had no official singles nor music videos for any of its songs), in October 2000 Kid A captured the number one spot on the Billboard 200 sales chart in its debut week. Unlike Metallica, Madonna and Dr.Dre that led litigation with Napster, Radiohead had never hit the top20 in the US before.
}
\cite{wiki_amnaptrial_2001}. This triggered a shift towards decentralised networks like the classic \emph{Gnutella} and, its ``rival'' at the time, \emph{FastTrack}. Gnutella, in its first realisation (v.0.4 \cite{gnutellav0.4}), was a fully decentralised approach. Unfortunatelly, network resourses tended to be overwhelmed by message flooding and if queries were constrained to live for some (small) number of hops, this ultimatelly resulted to a shrinked search scope, leading to a death spiral that prevented the network from scaling. The solution came with a slightly more structured network by introducing the notions of \emph{ultra-peers} and \emph{leaf-nodes} (v.0.6 \cite{gnutella}). With \emph{Limewire}, the most widely used software that implemented the protocol, the network was highly successful and was reported to have reached connecting more that 3 million users concurrently by the end of 2006 \cite{rsr_gnutellaevol_2006}. Fasttrack and its implementing software KaZaA \cite{kazaa}, on the other hand, went with the hierarchical approach (\emph{super-nodes} in their context) from the very begining, dynamically assigning indexing functions to special peers in order to increase search performace and scaling potential. FastTrack network had an average of 3 million concurrently connected nodes in 2003 \cite{lkr_kazaameasure_2005}. It is estimated that the total number of users was greater than that of Napster at its peak. With Limewire discontinued/shutdown in 2010 and its successor, \emph{Frostwire}, moving to the Bittorent protocol in 2011, seem to very well signal the end of the once-mighty Gnutella network. KaZaA continues as a propriatery music store solution. In the meanwhile, Fastrack design team had already introduced a VoIP and text messaging application based on the design principles of KaZaA's hierarchical P2P nature, that was meant to change how people communicate around the world. This app was no other than Skype \cite{skype}. Despite its wide usage, little is known for the technical aspects of its overlay network, mainly due to its proprietary nature. Studies, though, confirm that it shares strong structural similarities to its file-sharing predecessor (ordinary hosts and super-nodes in Skype's terminology) but they significantly differentiate in terms of traffic behaviour \cite{g_voipskypesec_2005, gdj_skypestudy_2006}. Skype was introduced in 2003 and from then on it witnessed frenzy-rythmed increase to its popularity, with more that 650 milion registered users in 2011  \cite{skypetotalusers} and reported to have broken the barrier of 35 million skypers concurrently online on the the 5th of March 2012 \cite{skypesymusers}. On October 2011 Microsoft Corp. and Skype Global S.à.r.l. anounnced the completion of Microsoft's acquisition of Skype for \$$8.5$ billion \cite{skypemicrosoft}.

File sharing was the most popular application area for structured architectures as well. Overnet
% TODO find citation
\cite{},
for example, was created for sharing large files (e.g., movies and CD images). It implemented the Kademlia tree-like partitioning algorithm. Other implementations of Kademlia are the Kad Network (eMule/aMule  clients), BitTorrent for trackerless torrernts (official BitTorrent client, $\mu$Torrent, BitComet, Transmission and BitSpirit all share compatibility with the Mainline DHT which realises the Kademlia protocol) and the Gnutella DHT (first implemented by LimeWire as the Mojito DHT). Lately, RetroShare \cite{retroshare} is the new trend for file-sharing which incorporates an implementation of the Kademlia DHT to create a Friend-to-Friend (F2F) network with some attributes of a darknet as well. But the potential of structured architectures was not exhausted only to the above. Structured P2P networks became popular in distributed systems because they could scale smoothly and have fast and reliable object location mechanisms. Thus, research and development was very active and productive resulting in a variety of diverse applications such like distributed search engines \cite{yaci}, distributed data storage systems \cite{kbc_oceanstore_2000, bdet_fsdfs_2000, dkkms_cfs_2001, dr_pastutility_2001, abc_farsite_2002, mmfc_ivy_2002, arla, agebh_dks_2003}, Web caches, archives, and publishing systems \cite{ird_squirrel_2002, bags_youserv_2002, wrc_publius_2000, wm_tangler_2001}, messaging applications \cite{threedegrees}, event notification infrastructures \cite{rkcd_scribe_2001, cdkr_scribe_2002, agebh_dks_2003}, naming services \cite{cmm_chorddns_2002}, censor-resistent stores \cite{cswh_freenet_2001} and lately even cloud-based platforms \cite{mgpj_cloudsnap_2011}.

%%%%%%%% NEW TRENDS %%%%%%%%
%File sharing Limewire Kazzaa The file-sharing landscape is slowly adjusting in response to the continued push for more anti-piracy tools, the final Pirate Bay verdict, and the raids and arrests in the Megaupload case, SOPA, ACTA. This leads to users trying to find new solutions for anonymous, uncensored, decentralised file sharing solutions like Freenet, Tribler and RetroShare where connections are made only between trusted peers — sometimes called "friends" and the respective formed network F2F.

% TODO: READ Dalesa published at eAsia 2009, http://www.e-asia.org/2009/ET_Presentations/Wathsala.pdf

The huge numbers of users, for the various applications we presented above, reveals the tremendous popularity P2P architectures have enjoyed. This, though, does not come without any toll. Measurements on some popular systems, such as FastTrack, Gnutella and DirectConnect have shown that the P2P traffic contributes the largest portion of the overall Internet traffic \cite{seroiu_analysiscds_2002, sen_analyzep2ptraffic_2004}. Even with later studies, this observation still holds on. Using deep packet inspection techniques deployed at selected customer sites \emph{Ipoque} and \emph{CacheLogic} were able to track down application usage \cite{cachelogic, ipoque2007, ipoque2009}. CacheLogic reports that more than 60\% of the traffic is identified as P2P, with Ipoque verifying this observation reporting similar results for 2007, while observing some small decline for 2008/2009 (but still remaining dominant compared to other applications). Apart from its participation in overall usage terms, P2P is already a huge contibutor to the overall data exchanged across the Internet\footnote{
    Mainly because file-sharing P2P is used to exchange big files of HD video or high sampled audio content compared to simple Web content distributed via, for example, HTTP.
}.
\cite{multinteligence} claims that P2P traffic in 2007 was around 1.6 petabytes per month and the potential is to be quintubled by the end of 2012 reaching almost 8 petabytes per month.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{The Topology Mismatch Problem}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
One of the major issues that defines the efficiency of an overlay network is the mapping of the overlay structure to the underlying infrastructure. Remember, in Figure~\ref{figure:overlay} Nodes $A$ and $B$ are on the same local network, while $C$ and $D$ are in different networks. The top layer represents the overlay interconnection formed by these nodes at a higher level. Link connections there, can change as needed by the running environment, without any particular constraint by the underlying physical topology. Let's have a closer look with the help of a simple example. Assume nodes $A$, $B$, $C$ and $D$ are connected through the physical network shown in Figure~\ref{figure:phys}, where the network costs are given in milliseconds, and peer $A$ sends a message to peer $D$.

If these peers participate in an overlay network according to one of the setups of Figure~\ref{figure:overlay-confs} then users will expierience different performances. In the overlay depicted in Figure~\ref{figure:overlay-1}, the message will traverse the following sequence of links in the physical layer (marked with their costs): $3 \rightarrow 7 \rightarrow 1 \rightarrow 2 \rightarrow 2 \rightarrow 1 \rightarrow 7 \rightarrow 8 \rightarrow 5 \rightarrow 5 \rightarrow 4$. In the alternative overlay of Figure ~\ref{figure:overlay-2}, the path will be: $3 \rightarrow 8 \rightarrow 4$. The total cost of the first sequnce is $45 ms$, while the second yields a mere $15 ms$. Therefore, we can conclude that the second overlay is more congruent with the underlying physical network than the first one and, thus, is more efficient.

\begin{figure}[ht]
\centering
\subfigure[Nodes B and C have direct overlay connection.] {
  \includegraphics[scale=0.8]{img/overlay-1.pdf}
  \label{figure:overlay-1}
}\qquad\qquad
\subfigure[Nodes A and D have direct overlay connection.] {
  \includegraphics[scale=0.8]{img/overlay-2.pdf}
  \label{figure:overlay-2}
}
\caption{Two different overlay connection configuration.}
\label{figure:overlay-confs}
\end{figure}

Ideally, overlay networks are expected to achieve an optimal mapping among the overlay paths and the underlying physical links, and avoid inefficient states. The problem of constructing an optimal overlay is referred to as the \emph{topology mismatch problem}, and formally defined as follows:
\begin{definition}
    Let $V = \{v_1, ..., v_n\}$ be a set of points denoting the network nodes, $\{v_i, v_j\} \in E$ be the set of unicast distances between nodes $v_i$ and $v_j$, $G=(V,E)$ be a complete distance graph over $V$. The topology mismatch problem is to construct a minimal spanning tree,  where node degree is restricted to a constant ($k\geq 2$) by the bandwidth of each node $v_i$.
\end{definition}

Early incarnations of the overlay protocols, however, did not make use of the optimal mapping with the physical network topology. Gnutella protocol for example was considered far from scalable \cite{ritter_gnucantscale_2001}. Since every peer can chοse its logical neighbours without knowledge of the underlying network had as a result a mismatch between physical and application-level topologies. Additionally, queries may be flooded to multiple paths merging in one peer while
in such case one of the paths would have been enough. Nevertheless, peers may forward the same message to each other before they receive the query messages from each other. \cite{matei_mapgnutella_2002} showed that, even given that 95\% of any two nodes are less than 7 hops away from each other, a flooding-based algorithm, like the one used by Gnutella, can generate 330TB/month in a 50.000 node network! Therefore, P2P traffic can have a serious effect on the overall well-being of the Internet as a whole.
